### 2020-01-11 paper/paper.tex
intro

Here we propose to simplify and improve calibration programs for
\eprv\ hardware systems with two very simple but novel ideas.
The first flows from the observation that calibration sources---which
include lamps, etalons, and laser-frequency combs (\lfc
s)---illuminate the spectrograph with dense sets of lines; almost every
location in the spectrograph image plane is surrounded by nearby,
useful calibration lines.
This recommends a calibration methodology that is
\emph{non-parametric}:
If every point in the spectrograph detector is surrounded by nearby
calibration lines, the wavelength solution can be made simply an
interpolation of the calibration data.
There is no need to choose any functional form for the wavelength
solution, such as an eighth-order polynomial.
Going non-parametric will improve calibration accuracy, because
choosing a parametric form biases the calibration, and biases it
badly when the form is inappropriate (as, for example, polynomials
will be at the detector edges).

The second idea here flows from the observation that contemporary
\eprv\ instruments are incredibly stable.
Temperature-controlled, fiber-fed pectrographs vary only slightly over
the night or season, and only along a small number of axes in what you
might call ``calibration space''..... HOGG

HOGG: Return to the point that non-parametric \emph{increases} the
freedom of the model (infinitely in some sense) but hierarchical
\emph{restricts} the freedom of the model (enormously, but we hope
appropriately).... HOGG

>>>>

Here we propose to simplify and improve calibration programs for
\eprv\ hardware systems with two very simple but novel ideas.
The first flows from the observation that calibration sources---which
include lamps, etalons, and laser-frequency combs (\lfc
s)---illuminate the spectrograph with dense sets of lines; almost every
location in the spectrograph image plane is surrounded by nearby,
useful calibration lines.
This recommends a calibration methodology that is
\emph{non-parametric}:
If every point in the spectrograph detector is surrounded by nearby
calibration lines, the wavelength solution can be made simply an
interpolation of the calibration data.
There is no need to choose any functional form for the wavelength
solution (such as an eighth-order polynomial, for example).
Going non-parametric will improve calibration accuracy, because
choosing a parametric form biases the calibration, and biases it
badly when the form is inappropriate (as, for example, polynomials
will be at the detector edges).

The second idea here flows from the observation that contemporary
\eprv\ instruments are incredibly stable.
Temperature-controlled, fiber-fed pectrographs vary only slightly over
the night or season, and only along a small number of axes in what you
might call ``calibration space'', or the (very high dimensional) space
of all possible wavelength solutions.
That is, not only are they stable, but they don't have many accessible
degrees of freedom.
So it is not appropriate to fit each calibration exposure or calibrate
each science exposure independently.
It makes sense to use the calibration data (or all the data) to
determine the space in which the instrument can and does vary, and
then in subsequent calibration work only permit the solution to live
in the accessible part of calibration space.
This structure is \emph{hierarchical}: The calibration data are used
not just to determine the wavlength solution, but also to determine
the possible space of wavelength solutions.
In the statistics literature this concept is often described as
\emph{de-noising}:
Every calibration exposure contains information about every other
exposure; it doesn't just constrain the instrument locally in time.
Thus every exposure can be improved (de-noised) with information
coming from every other exposure.

The method we propose here---\name---embodies these ideas.
It is a non-parametric, hierarchical, data-driven model for the
wavelength solution.
By being non-parametric, it delivers enormous freedom to the
wavelength solution to match or adapt to any instrument or detector
oddities.
By being hierarchical, it restricts that freedom tremendously, but it
does so appropriately for the observed variations in the
spectrography.
\name\ is designed for temperature-controlled, bench-mounted, fiber-fed
spectrographs with good calibration sources, such as laser-frequency
combs, etalons, or good arc lamps.
We have in mind the \eprv\ instruments and \eprv\ science cases.
But we expect \name\ to have applications for other kinds of
spectrographs in other contexts.
